---
title: "Lab 1"
author: "Alicia Fennell"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(tmap)
library(sf)
```

R markdown combines formatted text and code and outputs! Great for reproducibility - less opportunities for mistakes.

First we attach packages (in the code chunk above)

If you need to install package: go to the console, type install.packages("packagename")

Read in data (command option i for code chunk)
(command enter runs code)

# largest header
###### smallest header 

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

  
```

##Basic wrangling reminders

refresh data wrangling skills! 

Find top 5 highest observations of trees by legal status, then do some wrangling and make a graph. 

(command shift M is pipe operator)

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n= 5)

## count basically combines group by, n, summarize functions. super useful
##drop_na removes any rows that contain a missing or na value for the variable you specify 
# rename - new name goes first, then old name 
# relocate: tree_count moves to the first column 
#slice_max allows you to ID the rows with highest values for variable that you specify, and then only keeps top ## 
```

Make a graph of top 5 observations by legal status 

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count", title = "Test Title") +
  coord_flip() +
  theme_minimal()
  
```
A few more data wrangling refresher examples!!

Only want to keep observations (rows) for blackwood acacia trees
- no separate column for scientific and common names... but we can look for everything that contains "blackwood acacia" using filter. 

```{r}
blackwood <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood, aes(x = longitude, y = latitude)) +
  geom_point()


##string detect (str_detect) looks for a string within a variable that we specify 
# select helps us pick columns 
```
Use tidyr :: separate and unite functions. useful for combining or separating columns 

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_sci", "spp_common"), sep = "::")

```

Example of unite... (not sure why we'd do this!)

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "!!!!!")
```

Make some actual maps of Blackwood Acacia trees in SF. 

We'll use st_as_sf to convert lat and long values to spatial coordinates 

```{r}
blackwood_spatial <- blackwood %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_spatial) = 4326

ggplot(data = blackwood_spatial) +
  geom_sf(color = "darkgreen") +
  theme_minimal()

##geom_sf is for plotting spatial data in ggplot! once we've set the coordinate system. 
#but this is still hard to interpret... 

```

Read in sf roads! to make this map make more sense 

```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

##need these in the same coordinate system! there's already an existing crs for this so we'll use st_transform

st_transform(sf_map, 4362)

ggplot(data = sf_map) +
  geom_sf()
```

Now combine tree observations and roads map!

```{r}
ggplot() +
  geom_sf(data = sf_map, size = 0.1, color = "darkgray") +
  geom_sf(data = blackwood_spatial, size = 0.4, color = "darkgreen") + 
  theme_void() +
  labs(title = "Blackwood Acacias in San Francisco")
```

Let's make this interactive!! 

```{r}
tmap_mode("view")

tm_shape(blackwood_spatial) +
  tm_dots()
```

