---
title: "Lab 1"
author: "Alicia Fennell"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(tmap)
library(sf)
```

R markdown combines formatted text and code and outputs! Great for reproducibility - less opportunities for mistakes.

First we attach packages (in the code chunk above)

If you need to install package: go to the console, type install.packages("packagename")

Read in data (command option i for code chunk)
(command enter runs code)

# largest header
###### smallest header 

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

  
```

##Basic wrangling reminders

refresh data wrangling skills! 

Find top 5 highest observations of trees by legal status, then do some wrangling and make a graph. 

(command shift M is pipe operator)

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n= 5)

## count basically combines group by, n, summarize functions. super useful
##drop_na removes any rows that contain a missing or na value for the variable you specify 
# rename - new name goes first, then old name 
# relocate: tree_count moves to the first column 
#slice_max allows you to ID the rows with highest values for variable that you specify, and then only keeps top ## 
```

